Compositional equivalences based on Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-09-11 2016-Weak-FH-Bisim/WeakBisim.tex
qq commentaires en plus
correct lemma 8 + start thm 3bis

In
\cite{Johnson:2013:CBSE,Johnson:2014:qosa}, the authors investigate several methodologies for the compositional verification of software systems, with the aim to verify reconfigurable component systems. To improve scaling and compositionality, the authors  decouple the verification problem that is to be resolved by a SMT (satisfiability modulo theory) solver into independent sub-problems on independent sets of variables. These works clearly highlight the interest of incremental and compositional verification in a very general setting. By adding more structure to the composition model and designing a dedicated bisimulation theory, we show here how to enforce a compositional proof system that is more powerful than independent sets of variables. Our theory has also been encoded into an SMT solver and it would be interesting to investigate how the examples of evolving systems studied by the authors could be encoded into pNet and verified by our framework. However, the models of Johnson et al. are quite different from ours, in particular they are much less structured, and translating them is clearly outside the scope of this article. Additionally, the more compositional constraint resolution provided by the authors could also be interesting to make our SMT proofs scale better (inside a single pNet for example), the integration of the Z3 framework described in \cite{Johnson:2014:qosa} with our proof theory for open pNets would also be worth investigating from a practical point of view.

@@

\begin{proof} Suppose we have:
\begin{mathpar}
%\[
P\models{\openrule
	{
		\gamma_j^{j\in J}, 
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
		\ostate{s_i'^{\, i\in L}}}
}
\end{mathpar}
By Lemma \ref{lem-rel-OT-WOT} this implies the following: \\			
\begin{mathpar}
\forall p\!\in\![0..m_1].P\models\openrule
    {{\beta_{1pj}^{j\in J_{1p}}},\Pred_{1p},\Post_{1p}   }
         {\ostate{{(s_{1pi})}^{i\in L}} \OTarrow {\tau} \ostate{{(s_{1(p+1)i})}^{i\in L}}} \\
P\models\openrule
         {\beta_{2j}^{j\in J_2},\Pred_2,\Post_2 }
         {\ostate{({s_{20i}})^{i\in L}} \OTarrow {\alpha'} \ostate{({s_{21i}})^{i\in L}}}
\\
\forall p \in [0..m_2].P\models\openrule
         {
           \beta_{3pj}^{j\in J_{3p}},\Pred_{3p}, \Post_{3p}}
         { \ostate{(s_{3pi})^{i\in L}} \OTarrow {\tau} \ostate{(s_{3(p+1)i})^{i\in L}}}
\end{mathpar}
where for all $i\in L$:\\
$s_i=s_{10 i} \wedge s_{1(m_1+1) i}=s_{20i} \wedge  s_{21i}= s_{30 i} \wedge s_{3(m_2+1) i}=s'_i$\\
\[\alpha=\alpha'\displaystyle{\subst{\bigotimes_{j=m_1}^{0}\Post_{1j}}}\]
\begin{multline*}
\gamma_j^{j\in J}=\mybigdotcup_{i=0}^{m_1} \vis{{\beta_{1ij}^{j\in J_{1p}}}\subst{\bigotimes_{j=i-1}^{0} \Post_{1j} } }  \dotcup  \vis{{\beta_{2j}^{j\in J_{2}}}\subst{\bigotimes_{j=m_1}^{0}\Post_{1j}}} \\
\dotcup \mybigdotcup_{i=0}^{m_2} \vis{\beta_{3ij}^{j\in J_{1p}} \subst{\bigotimes_{j=i-1}^{0}\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=m_1}^{0}\Post_{1j}}}
\end{multline*}
\begin{multline*}
\Pred=\bigwedge_{p=0}^{m_1}\!\big(\Pred_{1p}\subst{\bigotimes_{j=p-1}^{0}\Post_{1j}}\big)\land\Pred^2\subst{\bigotimes_{p=m_1}^{0}\Post_{1p}}\land\\ \bigwedge_{p=0}^{m_2}\!\Pred_{3p}\subst{\bigotimes_{j=p-1}^{0}\Post_{3j}\otimes\Post_2 \otimes\bigotimes_{p=m_1}^{0}\Post_{1p}}
\end{multline*}
\begin{equation*}
\begin{split}
\Post&=\bigotimes_{p=m_2}^{0}\Post_{3p}\otimes\Post_2\otimes\bigotimes_{j=m_1}^{0}\Post_{1j}
\end{split}
\end{equation*}

Note that, for $l\in\{1,3\}$ if $\beta_{l p j_0}=\tau$, then, because of Definition \ref{def:Non-ObsTau}, $P$ necessarily makes a $\tau$ open transition and remains in the same state, ($s_p = s_{p+1}$). Thus without loss of generality, we can bypass such an open transition and obtain another decomposition of the WOT without the open transition that requires ${\beta^l_{p{j_0}}}=\tau$. We can thus  suppose that for all $p$ and $l$ we have ${\beta_{l p{j_0}}}\neq\tau$ or $j_0\not\in J_{1 p}$. To avoid a special case, we suppose that the hole $j_0$ moves during the OT $\alpha'$, i.e. $\beta_{2 \j_0}=\beta_{m}$ for some $m$. Additionally, $\beta_{m}\neq \tau$, else we would have $\alpha=\alpha'=\tau$ and the $\alpha'$ OT could be also removed from the reduction, leading to a particular and simpler case.

@@

  \newpage
\section{Examples of Lotos laws}
Possible Lotos small examples:

>>>>

\TODO{rertavailler le paragraphe ci dessous}
In
\cite{Johnson:2013:CBSE,Johnson:2014:qosa}, the authors investigate several methodologies for the compositional verification of software systems, with the aim to verify reconfigurable component systems. To improve scaling and compositionality, the authors  decouple the verification problem that is to be resolved by a SMT (satisfiability modulo theory) solver into independent sub-problems on independent sets of variables. These works clearly highlight the interest of incremental and compositional verification in a very general setting. By adding more structure to the composition model and designing a dedicated bisimulation theory, we show here how to enforce a compositional proof system that is more powerful than independent sets of variables. Our theory has also been encoded into an SMT solver and it would be interesting to investigate how the examples of evolving systems studied by the authors could be encoded into pNet and verified by our framework. However, the models of Johnson et al. are quite different from ours, in particular they are much less structured, and translating them is clearly outside the scope of this article. Additionally, the more compositional constraint resolution provided by the authors could also be interesting to make our SMT proofs scale better (inside a single pNet for example), the integration of the Z3 framework described in \cite{Johnson:2014:qosa} with our proof theory for open pNets would also be worth investigating from a practical point of view. One major different however is that pNets in their current state are not much adapted to probabilistic verification contrarily to \cite{Johnson:2014:qosa} but they are better suited to verify complex behavioural properties.

@@

\begin{proof} Suppose we have:
\begin{mathpar}
%\[
P\models{\openrule
	{
		\gamma_j^{j\in J}, 
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
		\ostate{s_i'^{\, i\in L}}}
}
\end{mathpar}
By Lemma \ref{lem-rel-OT-WOT} this implies the following: \\			
\begin{mathpar}
\forall p\!\in\![0..m_1].P\models\openrule
    {{\beta_{1pj}^{j\in J_{1p}}},\Pred_{1p},\Post_{1p}   }
         {\ostate{{(s_{1pi})}^{i\in L}} \OTarrow {\tau} \ostate{{(s_{1(p+1)i})}^{i\in L}}} \\
P\models\openrule
         {\beta_{2j}^{j\in J_2},\Pred_2,\Post_2 }
         {\ostate{({s_{20i}})^{i\in L}} \OTarrow {\alpha'} \ostate{({s_{21i}})^{i\in L}}}
\\
\forall p \in [0..m_2].P\models\openrule
         {
           \beta_{3pj}^{j\in J_{3p}},\Pred_{3p}, \Post_{3p}}
         { \ostate{(s_{3pi})^{i\in L}} \OTarrow {\tau} \ostate{(s_{3(p+1)i})^{i\in L}}}
\end{mathpar}
where for all $i\in L$:\\
$s_i=s_{10 i} \wedge s_{1(m_1+1) i}=s_{20i} \wedge  s_{21i}= s_{30 i} \wedge s_{3(m_2+1) i}=s'_i$\\
$\alpha=\alpha'\displaystyle{\subst{\bigotimes_{j=m_1}^{0}\Post_{1j}}}$\\
%\begin{multline*}
$\gamma_j^{j\in J}=\displaystyle{\mybigdotcup_{i=0}^{m_1} \vis{{\beta_{1ij}^{j\in J_{1p}}}\subst{\bigotimes_{j=i-1}^{0} \Post_{1j} } }  \dotcup \, \vis{{\beta_{2j}^{j\in J_{2}}}\subst{\bigotimes_{j=m_1}^{0}\Post_{1j}}}} \\
~~\qquad~\dotcup \mybigdotcup_{i=0}^{m_2} \vis{\beta_{3ij}^{j\in J_{3p}} \subst{\bigotimes_{j=i-1}^{0}\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=m_1}^{0}\Post_{1j}}}$\\
%\end{multline*}
%\begin{multline*}
$\Pred=\displaystyle{\bigwedge_{p=0}^{m_1}\!\big(\Pred_{1p}\subst{\bigotimes_{j=p-1}^{0}\Post_{1j}}\big)\land\Pred_2\subst{\bigotimes_{p=m_1}^{0}\Post_{1p}}\land}$\\ 
$~~\qquad\quad\displaystyle{\bigwedge_{p=0}^{m_2}\!\Pred_{3p}\subst{\bigotimes_{j=p-1}^{0}\Post_{3j}\otimes\Post_2 \otimes\bigotimes_{p=m_1}^{0}\Post_{1p}}}$\\
$\Post=\displaystyle{\bigotimes_{p=m_2}^{0}\Post_{3p}\otimes\Post_2\otimes\bigotimes_{j=m_1}^{0}\Post_{1j}}$
%\end{multline*}

Note that, for $l\in\{1,3\}$ if $\beta_{l p j_0}=\tau$, then, because of Definition \ref{def:Non-ObsTau}, $P$ necessarily makes a $\tau$ open transition and remains in the same state, ($s_p = s_{p+1}$). Thus without loss of generality, we can bypass such an open transition and obtain another decomposition of the WOT without the open transition that requires ${\beta_{lp{j_0}}}=\tau$. We can thus  suppose that for all $p$ and $l$ we have ${\beta_{lp{j_0}}}\neq\tau$ or $j_0\not\in J_{1 p}$. To avoid a special case, we suppose that the hole $j_0$ moves during the OT $\alpha'$, i.e. $\beta_{2 j_0}=\beta_{m}$ for some $m$. Additionally, $\beta_{m}\neq \tau$, else we would have $\alpha=\alpha'=\tau$ and the $\alpha'$ OT could be also removed from the reduction, leading to a particular and simpler case.

@@

\begin{theorem}[Congruence]\label{thm-congr-weq}
	Consider an open pNet:
	$\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\set{\symb{SV}}\myrangle$.
	Let $j_0\in J$ be a hole. Let $\pNetQ$ and $\pNetQ'$ be two weak FH-bisimilar pNets such that 
	$\Sortop(\pNetQ)=\Sortop(\pNetQ')=\Sort_{j_0}$. Then 
	$\pNet[\pNetQ]_{j_0}$ and 
	$\pNet[\pNetQ']_{j_0}$ are weak FH-bisimilar.
\end{theorem}
 \begin{proof}  (Sketch of proof) Consider $Q$ weak FH-bisimilar to $Q'$.  It means, according to Lemma  \ref{lem-alternative-weak-bisim} for any weak open transition $WOT$ in $Q$ there exist a family of weak open transitions $WOT_{x}^{x \in X}$ in $Q'$. 
We apply  Lemma \ref{lem-Weakcompose1} on weak open transition $WOT$, $j_0 \in J$, we obtain a new weak open transition $WOT'$ in $P[Q]_{j_0}$.
On other hand, we apply same Lemma \ref{lem-Weakcompose1} on each weak open transitions $WOT_{x}$, so we obtain a new family of weak open transitions $WOT_{x}'$ in $P[Q']_{j_0}$.
 \end{proof}
\begin{theorem}[Context equivalence]\label{thm-ctxt-weq}
	Consider two FH-bisimilar open pNets:
	$\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\set{\symb{SV}}\myrangle$ and 	$\pNet' = \mylangle {\pNet'}_i^{i\in I}, 
	\Sort_j^{j\in 
	J}, 	\set{\symb{SV'}}\myrangle$ 
	(recall they must have the same holes to be bisimilar).
	Let $j_0\in J$ be a hole, and $Q$ be a pNet such that $\Sortop(Q)=\Sort_{j_0}$. Then 
	$\pNet[Q]_{j_0}$ and 
	$\pNet'[Q]_{j_0}$ are FH-bisimilar.
\end{theorem}
  \newpage
\section{Examples of Lotos laws}
Possible Lotos small examples:
